### 2011-01-21 paper/learningMethods.tex
More learning methods work (mostly)

\subsection{Learning Methods}

\edit{write this.  Old quadratot Method section below}

%\section{Method}
\seclabel{method}

@@

We use several parameterized motion models that command motors to
positions based on a sine wave, creating a periodic pattern.  While we
investigated several models, for the bulk of our experiments, we used
a model whose five parameters are: amplitude, wavelength, scale inner
vs outer motors, scale left vs right motors, scale back vs front
motors. Each strategy below attempts to choose the best possible
parameters for this motion model.

We implemented and tested 8 different learning strategies.  All
strategies except for the HyperNEAT method\cite{clune} were
constrained to pick parameters from within predetermined ranges.

@@

\item \emph{Random}: This method randomly generates parameter vectors
  in the allowable range. This strategy is used only as baseline.

@@

\item \emph{N-dimensional policy gradient descent}: As opposed to the
  previous methods, this method explicitly estimates the gradient for
  the objective function. It does this by first evaluating \emph{t}
  randomly generated parameter vectors near the initial vector, each
  dimension of these vectors being perturbed by either $-\epsilon$,
  $0$, or $\epsilon$. Then, for each dimension, it groups vectors into
  three groups: $-\epsilon$, $0$, and $\epsilon$.  The gradient along
  this dimension is then estimated as the average score for the
  $\epsilon$ group minus the average score for the $-\epsilon$
  group. Finally, the method creates a new vector by changing all
  parameters by a fixed-size step in the direction of the gradient.

\item \emph{Nelder-Mead simplex method}\cite{nm}: The Nelder-Mead
  simplex method creates an initial simplex with 6 vertices. The
  initial parameter vector is stored as the first vertex and the other
  five vertices are created by adding to one dimension at a time one
  tenth of the allowable range for that parameter. It then tests the
  fitness of each vertex and based on these fitnesses, it reflects the
  worst point over the centroid in an attempt to improve it.  However,
  to prevent cycles and becoming stuck in local minima, several other
  rules are used.  In general, the worst vertex is reflected over the
  centroid. If the reflected point is better than the second worst
  point and worse than the best point, then the reflected point
  replaces the worst. If the reflected point is better than the best
  point, the simplex is expanded in the direction of the reflected
  point. The better of the reflected and the expanded point replaces
  the worst point. If the reflected point is worse than the second
  worst point, then the simplex is contracted away from the reflected
  point. If the contracted point is better than the reflected point,
  the contracted point replaces the worst point. If the contracted
  point is worse than the reflected point, the entire simplex is
  shrunk \cite{nm}.

>>>>

\subsection{Parametrized Motion Models}
\seclabel{motionModel}

\edit{this might no longer flow...fix} There are a plethora of methods available for the creation of a vector motion generating function $g(t)$, and we've chosen two for comparison in this study.

The first method is to choose a family of \emph{parametrized functions}.  By fixing the parameters to a particular set of values, we obtain a deterministic motion function over time.  We tried several parametrizations on the robot and, obtaining reasonble early success, settled on one particular parametrization.  We dubbed this the \emph{SineModel5}, as it has five parameters and employs a sine wave as its root pattern.

The five parameters in $\vec{\theta}$ are as follows:

 - amp - Amplitude       - allowable range
 - tau   - sine wave period   - allowable range
 - multIO - multiplier scaling inner joings from outer joints  - allowable range
 - multFB - multiplier scaling front joints from back joints  - allowable range
 - multLR - multiplier scaling left joints from right joints  - allowable range
(verify this)

\newcommand{\amp}{\mathrm{amp}}

Inuitively, SineModel5 starts with 8 identicaly sine waves of amplitude $\amp$ and period $\tau$, multiplies the waves for all inner motors by multIO, multiplies the waves for all front motors by multFB, and multiplies the waves for all left motors by multLR.  To obtain the actual motor position commands, these waves are offset by an appropriate fixed constant so that the base position (when the sine waves are at 0) is approximately a crouch (the position shown in \figref{crouchingRobot}).  Finally, in this motion model, the ninth "hip" motor is kept at a neutral position.  Thus, the commanded position for all motors, as a vector function of time, is:

\begin{verbatim}
        [g_0(t) ]    [ \amp * sin(\tau t) ]
        [g_1(t) ]    [ \amp * sin(\tau t) ]
        [g_2(t) ]    [ \amp * sin(\tau t) ]
g(t)   [g_3(t) ] = [ \amp * sin(\tau t) ]
        [g_4(t) ]    [ \amp * sin(\tau t) ]
        [g_5(t) ]    [ \amp * sin(\tau t) ]
        [g_6(t..) ]    [ \amp * sin(\tau t) ]
\end{verbatim}

The nine subscripted functions correspond to the nine labeled motors in \figref{robotDiagram}

% used is a function parametrized by five numbers, shown in \figref{foo}

\subsection{Learning Methods for Parametrized Motion Models}
\seclabel{learningMethods}

Using the SineModel5 parametrized motion model described in the previous section, along with the allowable ranges for each of the five parameters (shown in \tabref{parameterTable}), the task becomes how to choose the combination of five parameters that results in the fastest motion, per the evaluation methods in \secref{fitnessEvaluation}.

If we choose a value for the five dimensional parameter $\vec{\theta}$, then a given physical trial gives us one measurement of the fitness of that parameter vector, or $f(\vec{\theta})$.  Two things make learnign difficult.  First, each evaluation of $f(\vec{\theta})$ is expensive, taking 15-20 seconds of time on average.  Second, the fitness returned by such evaluations has proven to be very noisy, with the standard deviation of the noise often being roughly equivalent to the size of the measurement.

For both of these reasons, an intelligent learning algorithm is needed.

In this context, by learning algorithm we simply mean a method for choosing:

1) The value for $\vec{\theta}$ for the initial trial, and
2) The next value of $\vec{\theta}$ to try, given the $\vec{\theta}$ and $f(\vec{\theta})$ for all previous trials

We evaluated a total of eight subtly or not so subtly different learning methods.  All employed a simple random sampling method for requirement (1); that is, all methods picked their initial $\vec{\theta}$ value via uniform random sampling within the allowed parameter ranges. \edit{make sure this is true}.  Thus, the differences in the algorithms was how the selected new $\vec{\theta}$ values to try from their past experience.

The eight methods used, and their methods for choosing the next $\vec{\theta}$ are:

\edit{write this.  Old quadratot Method section below alsdkfj }

%\section{Method}
%\seclabel{method}

@@

%We use several parametrized motion models that command motors to
%positions based on a sine wave, creating a periodic pattern.  While we
%investigated several models, for the bulk of our experiments, we used
%a model whose five parameters are: amplitude, wavelength, scale inner
%vs outer motors, scale left vs right motors, scale back vs front
%motors. Each strategy below attempts to choose the best possible
%parameters for this motion model.

%We implemented and tested 8 different learning strategies.  All
%strategies except for the HyperNEAT method\cite{clune} were
%constrained to pick parameters from within predetermined ranges.

@@

\item \emph{Random}: This method randomly generates parameter vectors
  in the allowable range for every trial. This strategy was used only as baseline for comparison.

@@

\item \emph{N-dimensional policy gradient descent}: In \cite{kohl},
  Kohl and Stone document a method for local gradient ascent for gait
  learning with noisy fitness evaluations, and we have included this
  method in our evaluation.  This strategy explicitly estimates the
  gradient for the objective function. It does this by first
  evaluating \emph{n} randomly generated parameter vectors near the
  initial vector, each dimension of these vectors being perturbed by
  either $-\epsilon$, $0$, or $\epsilon$. Then, for each dimension, it
  groups vectors into three groups: $-\epsilon$, $0$, and $\epsilon$.
  The gradient along this dimension is then estimated as the average
  score for the $\epsilon$ group minus the average score for the
  $-\epsilon$ group. Finally, the method creates a new vector by
  changing all parameters by a fixed-size step in the direction of the
  gradient.

\item \emph{Nelder-Mead simplex method}: The Nelder-Mead simplex
  method \cite{nm} creates an initial simplex with $d+1$ vertices,
  where $d$ is the dimension of the parameter space. The initial
  parameter vector is stored as the first vertex and the other five
  vertices are created by adding to one dimension at a time one tenth
  of the allowable range for that parameter. It then tests the fitness
  of each vertex and based on these fitnesses, it reflects the worst
  point over the centroid in an attempt to improve it.  In general,
  the worst vertex is reflected; however, to prevent cycles and
  becoming stuck in local minima, several other rules are used.  If
  the reflected point is better than the second worst point and worse
  than the best point, then the reflected point replaces the worst. If
  the reflected point is better than the best point, the simplex is
  expanded in the direction of the reflected point. The better of the
  reflected and the expanded point replaces the worst point. If the
  reflected point is worse than the second worst point, then the
  simplex is contracted away from the reflected point. If the
  contracted point is better than the reflected point, the contracted
  point replaces the worst point. If the contracted point is worse
  than the reflected point, the entire simplex is shrunk \cite{nm}.
